---
title: "Practical Machine Learning Project"
author: "KuangyuWang"
date: "04/20/2015"
output: html_document
---

# Data preprocessing

### Load libraries

```{r message=FALSE}
library(RCurl)
library(R.utils)
```

### Download training and testing data files

```{r}
# Download training data
if(!file.exists("~/workspace/pml_proj/data/pml-training.csv")){
    download.file("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", "~/workspace/pml_proj/data/pml-training.csv", method="auto")
}
# Download testing data
if(!file.exists("~/workspace/pml_proj/data/pml-testing.csv")){
    download.file("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", "~/workspace/pml_proj/data/pml-testing.csv", method="auto")
}
```

### Read in data

```{r cache=TRUE}
trainingDT = read.csv("~/workspace/pml_proj/data/pml-training.csv")
testingDT = read.csv("~/workspace/pml_proj/data/pml-testing.csv")
```

### Inspect features

Remove numeric variables with too much missing data or have *near zero variance*

```{r}
allNumericVars = colnames(trainingDT)[sapply(trainingDT, is.numeric)]
numeric2keepVars = allNumericVars

# find vars
```

# Model building



built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases. 
